---
title: "Simple Linear Regression Incomne/Happiness"
author: "Karl Mbouombouo"
date: "2023-08-21"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(broom)
library(ggpubr)
```

## Data
```{r}
income_data <- read_csv("income.data.csv")
summary(income_data)
```

## Plots
```{r}
hist(income_data$happiness)
```
The observations are "bell-shaped", this is a good to proceed a linear model

```{r happiness}
plot(happiness ~ income, data = income_data)
```

The relationship between income and happiness look pretty linear, so we can
use the linear model

## Modeling
```{r}
income.lm <- lm(happiness ~ income, data = income_data)
summary(income.lm)
```
From these results, we can say that there is a significant positive relationship
between income and happiness (p value < 0.001), with a 0.71383 unit increase in 
happiness for every unit increase in income.

## Check Homoscedasticity
```{r}
par(mfrow=c(2,2))
plot(income.lm)
par(mfrow=c(1,1))
```
The residuals are for most of them all horizontal and centered to zero, this is 
mean there is no biases or outliers in the data that would make the linear
regression invalid

## Visualization of result
```{r}
income.graph<-ggplot(income_data, aes(x=income, y=happiness))+
                     geom_point() +
                     geom_smooth(method="lm", col="red") +
                     stat_regline_equation(label.x = 3, label.y = 7) +
                     theme_bw() +
                     labs(title = "Reported happiness as a function of income",
                     x = "Income (x$10,000)",
                     y = "Happiness score (0 to 10)")
  

income.graph
```

## Conclusion
There is a significant relationship between income and happiness (p < 0.001, 
R2 = 0.73 ± 0.0193), with a 0.73 unit increase in reported happiness for every 
$10,000 increase in income.