---
title: "Perfomance Task"
author: "Karl Mbouombouo"
date: "2023-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r, include=FALSE}
library('tibble')
library("tidyr")
library("dplyr")
library("ggplot2")
library("readr")
library("glmnet")
library("pscl")
library("caret")
library("remotes")
library("InformationValue", "1.2.3")
library("plotROC")
library("pROC")
library("hrbrthemes")
```

## Import, and "clean" the data

```{r, include=FALSE}
mydata <- read_csv("Documents/PerfomTask/Analyst_Performance_Tasks_Data_Set - Analyst_Performance_Tasks_Data_Set.csv")
which(is.na(mydata))

df_gradelevel <- mydata[mydata$grade_level == '9',]
df_markingperiode <- mydata[mydata$marking_period == 'Fall',]
df_course <- mydata[mydata$course == 'Math 9',]
#Remove columns "marking_period", "school_code", "sid", "grade_level" and "course" because the school_id is not needed since we have school_name, neither sid because it's just student number id, we can remove grade_level, course since they are all for grade 9 students, and we can remove marking_period since they are all during the Fall. School A Sample too small to considered. This is reduce let us with 8 schools instead of 9. Graph percent of success per school
df = subset(mydata, select = -c(school_year, school_code, sid, grade_level, marking_period, course))
df<- df[!(df$school_name %in% c("SCHOOL A")),]

#Raking for grade, A = B = 1, C = 2, D = 3 and F=I=4
df2 <- df %>% mutate(ranking = case_when(
          (df$final_grade_mark == "A") ~ "1",
          (df$final_grade_mark == "A-") ~ "1",
          (df$final_grade_mark == "A+") ~ "1",
          (df$final_grade_mark == "B") ~ "1",
          (df$final_grade_mark == "B-") ~ "1",
          (df$final_grade_mark == "B+") ~ "1",
          (df$final_grade_mark == "C") ~ "0",
          (df$final_grade_mark == "C-") ~ "0",
          (df$final_grade_mark == "C+") ~ "0",
          (df$final_grade_mark == "D") ~ "0",
          (df$final_grade_mark == "D-") ~ "0",
          (df$final_grade_mark == "D+") ~ "0",
          (df$final_grade_mark == "F") ~ "0",
          (df$final_grade_mark == "P") ~ "1",
          (df$final_grade_mark == "I") ~ "0"
          ))

#Since we are looking for 1-3 school with the best program, we will create a new column with "Success" with the grades greater than F, 1 is for success passing and 0 for failed passing.
df3 <- df2 %>% mutate(success = case_when(
          (df$final_grade_mark == "A") ~ "Yes",
          (df$final_grade_mark == "A-") ~ "Yes",
          (df$final_grade_mark == "A+") ~ "Yes",
          (df$final_grade_mark == "B") ~ "Yes",
          (df$final_grade_mark == "B-") ~ "Yes",
          (df$final_grade_mark == "B+") ~ "Yes",
          (df$final_grade_mark == "C") ~ "Yes",
          (df$final_grade_mark == "C-") ~ "Yes",
          (df$final_grade_mark == "C+") ~ "Yes",
          (df$final_grade_mark == "D") ~ "Yes",
          (df$final_grade_mark == "D-") ~ "Yes",
          (df$final_grade_mark == "D+") ~ "Yes",
          (df$final_grade_mark == "F") ~ "No",
          (df$final_grade_mark == "P") ~ "Yes",
          (df$final_grade_mark == "I") ~ "No"
          ))
#Consider Former English learner as Never English learner, since they has gained enough English language
df4 <- df3 %>% mutate(el_or_not= case_when(
          (df$el_status == "Current EL") ~ "1",
          (df$el_status == "Former EL") ~ "0",
          (df$el_status == "Never EL") ~ "0"
          ))

#Put variable as factor
df4$success <- as.factor(df4$success) # Passing class = 1, Failing class =2
df4$ranking <- as.factor(df4$ranking)
df4$iep <- as.factor(df4$iep) # Don't Have a EP = 1, Have a EP = 2
df4$el_status <- as.factor(df4$el_status) # Current EL = 1, Former EL = 2, Never EL = 3
df4$ecodis <- as.factor(df4$ecodis) # not economically disadvantaged = 1, economically disadvantaged = 2
str(df4)

#Count of student per school
survey_by_school <- group_by(df2, school_name)
summarize(survey_by_school, counts = n())
#Count of student that passed their classes per school, passed = TRUE otherwise = FALSE
survey_by_school_count <- group_by(df3, school_name, success == "1")
summarize(survey_by_school_count, counts = n())
#Count of student with iep that passed their classes per school, passed = TRUE otherwise = FALSE
survey_by_school_iep <- group_by(df3, school_name, iep == "1")
summarize(survey_by_school_iep, counts = n())
```

# **ETL**

## Top 3 schools with the most students passing their course
```{r, include=FALSE}
#Percent of student that passed their course per school
Success_SchoolB = 471/578
Success_SchoolC = 691/753
Success_SchoolD = 481/562
Success_SchoolE = 519/587
Success_SchoolF = 805/867
Success_SchoolG = 532/635
Success_SchoolH = 670/814
Success_SchoolI = 829/880
Success_SchoolJ = 496/664
```

```{r, echo=FALSE}
#Pie Chart to show it
slices <- c(81, 75, 94, 83, 84, 93, 88, 86, 92)
labels <- c("School B", "School  J", "School I", "School H", "School G",
          "School F", "School E", "School D", "School C")
labels <- paste(labels, slices)
labels <- paste(labels,"%",sep="")
pie(slices,labels = labels, col=rainbow(length(labels)),
   main="Pie Chart of passing students in each school")
```

The Top 3 schools with the most students that passe their course are schools I with 94%, school F with 93% and school C with 92% Overall ranking: 1- School I: 94% 2- School F: 93% 3- School C: 92% 4- School E: 88% 5- School D: 86% 6- School G: 84% 7- School H: 83% 8- School B: 81% 9- School J: 75%

```{r, include=FALSE}
#Percent of student with iep that passed their course
iep_B = 127/578
iep_C = 138/753
iep_D = 76/562
iep_E = 100/587
iep_F = 168/867
iep_G = 115/635
iep_H = 168/670
iep_I = 133/880
iep_J = 144/664
```
```{r}
slices3 <- c(22, 22, 15, 25, 18, 19, 18, 14, 18)
labels <- c("School B", "School  J", "School I", "School H", "School G",
          "School F", "School E", "School D", "School C")
labels <- paste(labels, slices3)
labels <- paste(labels,"%",sep="")
pie(slices3,labels = labels, col=rainbow(length(labels)),
   main="Pie Chart of students with iep that pass their course")
```

## Top 3 schools with the most students passing their course with B- or higher

```{r, include=FALSE}
#Count of student that passed their classes per school, passed = TRUE otherwise = FALSE
survey_by_school_count <- group_by(df3, school_name, ranking == "1")
summarize(survey_by_school_count, counts = n())
#Percent of student that passed their course with a B- or higher per school
Top_SchoolB = 270/471
Top_SchoolC = 452/691
Top_SchoolD = 246/481
Top_SchoolE = 299/519
Top_SchoolF = 513/805
Top_SchoolG = 296/532
Top_SchoolH = 383/814
Top_SchoolI = 612/829
Top_SchoolJ = 270/496
```
```{r, echo=FALSE}
# Pie Chart to show it
slices2 <- c(57, 54, 74, 57, 56, 64, 58, 51, 65)
labels <- c("School B", "School  J", "School I", "School H", "School G",
          "School F", "School E", "School D", "School C")
labels <- paste(labels, slices2)
labels <- paste(labels,"%",sep="")
pie(slices2,labels = labels, col=rainbow(length(labels)),
   main="Pie Chart of students that their classes with B- or Higher")
```
The Top 3 schools with the most students that passe their course with B- or higher are schools I with 74%, school C with 65% and school F with 64% Overall ranking: 1- School I 74% 2- School C 65% 3- School F 64% 4- School E 58% 5- School B 57% 6- School H 57% 7- School G 56% 8- School J 54% 9- School D 51%

```{r}
survey_by_school_ieptg <- group_by(df3, school_name, iep == '1', ranking == "1")
print(summarize(survey_by_school_ieptg, counts = n()), n=36)
#Percent of student with iep that passed their course with a B- or higher per school
ieptg_B = 66/471
ieptg_C = 77/691
ieptg_D = 32/481
ieptg_E = 53/519
ieptg_F = 99/805
ieptg_G = 61/532
ieptg_H = 84/814
ieptg_I = 106/829
ieptg_J = 75/496
```
```{r}
slices3 <- c(14, 15, 13, 10, 11, 12, 10, 7, 11)
labels <- c("School B", "School  J", "School I", "School H", "School G",
          "School F", "School E", "School D", "School C")
labels <- paste(labels, slices3)
labels <- paste(labels,"%",sep="")
pie(slices3,labels = labels, col=rainbow(length(labels)),
   main="Pie Chart of students with iep that their classes with B- or Higher")
```


## Analyse iep, el_or_not and ecodis

```{r}
ggplot(data = df4, aes(x = iep, fill = school_name)) + 
  geom_bar(position = position_dodge())

ggplot(data = df4, aes(x = el_or_not, fill = school_name)) + 
  geom_bar(position = position_dodge())

ggplot(data = df4, aes(x = ecodis, fill = school_name)) + 
  geom_bar(position = position_dodge())
```

1- The schools that have the most individualized education plan installed are school H, F, J, C. School F and C are the schools with the highest percentage with students passing their classes with B- or higher. So, it might not be in relationship with the final grade because they do not have much students with disabilities than others. So we might considerate the effect that have individualized education plan on the final grades

2- The schools with the most student who have never been considered an English Learner are school I, F, C. Those are the school that have the most percentage with student that finish their course with a final great with a B- or higher. So we might exist a relationship between the students that never been considerate as English learning and their final grades.

3- The schools with the most student classified as economically disadvantaged are schools I, F, H. Once again school I and F are the second that have the best results regarding their final grades results. So, a relationship might exist between student budget and their final grades.

```{r}
ggplot(df3, aes(x=final_grade_mark, y=days_enrolled_at_school, color=school_name)) + 
    geom_point(size=2) +
    theme_ipsum()

ggplot(data = df3, aes(x = school_name, fill = days_suspended_out_of_school > 0)) + 
  geom_bar(position = position_dodge())

ggplot(data = df3, aes(x = school_name, fill = days_absent_at_school > 0)) + 
  geom_bar(position = position_dodge())
```

The top 3 schools that have the highest percentage with students that finish their course with a B- or higher are school I, C and F. But also those school are the school that have the most students passing their classes. But those school got a bigger sample of students, so they have more students to study and might affect the results of students with final grade with a B- or higher.

## Machine Learning - Logistics regression

```{r}
#Put variable as factor
df4$success <- as.factor(df4$success) # Passing class = 1, Failing class =2
df4$ranking <- as.factor(df4$ranking)
df4$iep <- as.factor(df4$iep) # Don't Have a EP = 1, Have a EP = 2
df4$el_status <- as.factor(df4$el_or_not) # Current EL = 1, Former EL = 2, Never EL = 3
df4$ecodis <- as.factor(df4$ecodis) # not economically disadvantaged = 1, economically disadvantaged = 2
str(df4)
```

```{r}
#Set the seed
set.seed(1234567)

#Use 70% of dataset as training set and remaining 30% as testing set
sample <- sample(c(TRUE, FALSE), nrow(df4), replace=TRUE, prob=c(0.7,0.3))
train <- df4[sample, ]
test <- df4[!sample, ] 

#Built the predictive model
model <- glm(success~iep+el_or_not+ecodis+days_enrolled_at_school+days_absent_at_school+
               days_suspended_out_of_school,
               family="binomial", data=train)
summary(model)
pscl::pR2(model)["McFadden"]
```

```{r}
plot(model)

caret::varImp(model)

# calculate probability of default for each individual in test dataset
predicted <- predict(model, test, type="response")

#convert success from "passed" and "failed" to 1's and 0's
test$success <- ifelse(test$success=="Yes", 1, 0)

#find optimal cutoff probability to use to maximize accuracy
optimal <- optimalCutoff(test$success, predicted)[1]
optimal

confusionMatrix(test$success, predicted)

#calculate sensitivity
sensitivity(test$success, predicted)

#calculate specificity
specificity(test$success, predicted)
```

#Misclassification

```{r}
#calculate total misclassification error rate
misClassError(test$success, predicted, threshold=optimal)
```

#Roc Curve

```{r}
#plot the ROC curve
plotROC(test$success, predicted)
```

We can see that the AUC is 0.8437, which is quite high. This indicates that our model does a good job of predicting whether or not an individual will pass the course.