---
title: "Basketball Analysis - NBA 2017-2018"
author: "Karl Mbouombouo"
date: "2023-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PACKAGES
```{r}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(vcd)
library(MASS)
library(lsr)
library(tibble)
library(base)
library(GGally)
library(tidygraph)
library(igraph)
library(dendextend)
```

```{r}
RNGkind(sample.kind="Rounding")

rm(list=ls())
# install.packages("devtools", repos="https://cran.stat.unipd.it/")
# devtools::install_github("sndmrc/BasketballAnalyzeR",force=TRUE)
library(BasketballAnalyzeR)

data(package="BasketballAnalyzeR")
nba_pbp <- PbPmanipulation(PbP.BDB)
```

# STATISTICAL ANALYSIS
```{r}
# Analysis TOP 6 Eastern Conference 
# four factor: Shooting actor, turnovers, rebounding, and free throws
rm(list=ls())

tm <- c("MIL","BOS","PHI","CLE", "NY", "BKN")
selTeams <- which(Tadd$team %in% tm)
FF.sel <- fourfactors(Tbox[selTeams,], Obox[selTeams,])

plot(FF.sel)
```

## Bar Plots
```{r}
X <- data.frame(Tbox, PTS.O=Obox$PTS, TOV.O=Obox$TOV,
                CONF=Tadd$Conference)
XE <- subset(X, CONF=="E")
labs <- c("Steals","Blocks","Defensive Rebounds")
barline(data=XE, id="Team", bars=c("STL","BLK","DREB"),
        line="TOV.O", order.by="PTS.O", labels.bars=labs)

ggplot(XE, aes(x=Team, y=TOV.O)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=TOV.O), vjust=-0.3, size=3.5) +
  ggtitle("Team Eastern Conference, Total Turn Over") +
  theme_minimal()

ggplot(XE, aes(x=Team, y=STL)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=STL), vjust=-0.3, size=3.5) +
  ggtitle("Team Eastern Conference, Total Steals") +
  theme_minimal()

ggplot(XE, aes(x=Team, y=BLK)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=BLK), vjust=-0.3, size=3.5) +
  ggtitle("Team Eastern Conference, Total Blocks") +
  theme_minimal()

ggplot(XE, aes(x=Team, y=DREB)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=DREB), vjust=-0.3, size=3.5) +
  ggtitle("Team Eastern Conference, Total Defensive Rebound") +
  theme_minimal()

```

```{r}
Pbox.WW <- subset(Pbox, Team=="Washington Wizards" &
                  MIN>=500)

barline(data=Pbox.WW, id="Player",
        bars=c("P2p","P3p","FTp"), line="MIN",
        order.by="PM", labels.bars=c("2P%","3P%","FT%"),
        title="Washington Wizards")

ggplot(Pbox.WW, aes(x=Player, y=MIN)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=MIN), vjust=-0.3, size=3.5) +
  ggtitle("Washington Wizard, Total Minutes") +
  theme_minimal()

ggplot(Pbox.WW, aes(x=Player, y=P2p)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=P2p), vjust=-0.3, size=3.5) +
  ggtitle("Washington Wizard, Total FG") +
  theme_minimal()

ggplot(Pbox.WW, aes(x=Player, y=P3p)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=P3p), vjust=-0.3, size=3.5) +
  ggtitle("Washington Wizard, Total 3FG") +
  theme_minimal()

ggplot(Pbox.WW, aes(x=Player, y=P2p)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=FTp), vjust=-0.3, size=3.5) +
  ggtitle("Washington Wizard, Total FT") +
  theme_minimal()
```

## Radial Plots
```{r}
attach(Pbox.WW)

X2 <- data.frame(P2M, P3M, FTM, REB=OREB+DREB, AST,
                STL, BLK)/MIN

detach(Pbox.WW)

# Standardization OFF
radialprofile(X2, title=Pbox.WW$Player, std=FALSE)
```

```{r pressure, echo=F}
# Standardization ON
radialprofile(X2, title=Pbox.WW$Player, std=TRUE)
```

## Scatter Plot
```{r}
Pbox.sel <- subset(Pbox, MIN>= 500)
attach(Pbox.sel)
X3 <- data.frame(AST, TOV, PTS)/MIN
detach(Pbox.sel)
mypal <- colorRampPalette(c("blue","yellow","red"))
scatterplot(X3, data.var=c("AST","TOV"), z.var="PTS",
            labels=Pbox.sel$Player, palette=mypal)
```

```{r}
WAS <- which(Pbox.sel$Team == "Washington Wizards")
scatterplot(X3, data.var = c("AST","TOV"), z.var="PTS",
            labels = Pbox.sel$Player, palette=mypal,
            subset = WAS)
```

```{r}
# Zoom the previous scatter plot
scatterplot(X3, data.var=c("AST","TOV"), z.var="PTS",
            labels=Pbox.sel$Player, palette=mypal,
            subset=WAS, zoom=c(0,0.25,0.05,0.10))
```

## Bubble Plot
```{r}
attach(Tbox)
X4 <- data.frame(T=Team, P2p, P3p, FTp, AS=P2A+P3A+FTA)
detach(Tbox)
labs <- c("2-point shots (% made)",
          "3-point shots (% made)",
          "free throws (% made)",
          "Total shots attempted")
bubbleplot(X4, id="T", x="P2p", y="P3p", col="FTp", size="AS", labels=labs)
```

```{r}
# Defensives stats: WAS vs. TOR
Pbox.WAS.TOR <- subset(Pbox,
               (Team=="Washington Wizards" |
                      Team =="Toronto Raptors") &
                      MIN>=500)
attach(Pbox.WAS.TOR)
WAS.TOR <- data.frame(ID=Player, Team, V1=DREB/MIN, V2=STL/MIN,
                V3=BLK/MIN, V4=MIN)
detach(Pbox.WAS.TOR)
labs <- c("Defensive Rebounds","Steals","Blocks",
          "Total minutes played")
bubbleplot(WAS.TOR, id="ID", x="V1", y="V2", col="V3",
           size="V4", text.col="Team", labels=labs,
           title="WAS and TOR during the regular season",
           text.legend=TRUE, text.size=3.5, scale=FALSE)
```

## Variability analysis
```{r}
vrb1 <- variability(data=Pbox.WW, data.var="P3p",
                    size.var="P3A")

vrb1 <- variability(data=Pbox.WW, data.var="P3p",
                    size.var="P3A",weight=TRUE)

vrb2 <- variability(data=Pbox.WW,
                    data.var=c("P2p","P3p","FTp"),
                    size.var=c("P2A","P3A","FTA"),
                    weight=TRUE)

plot(vrb2, title="Variability diagram - WAS")
```

## Inequality analysis
```{r}
ineqWW <- inequality(Pbox.WW$PTS, nplayers=8)

Pbox.BC <- subset(Pbox, Team == "Boston Celtics")
ineqBC <- inequality(Pbox.BC$PTS, nplayers=8)

p1 <- plot(ineqWW, title = "Washington Wizards")

p2 <- plot(ineqBC, title = "Boston Celtics")

grid.arrange(p1, p2, nrow = 1)
```

```{r}
no.teams <- nrow(Tbox)
INEQ <- array(0, no.teams)
for (k in 1:no.teams) {
     Teamk <- Tbox$Team[k]
     Pbox.sel <- subset(Pbox, Team==Teamk)
     index <- inequality(Pbox.sel$PTS, npl=8)
     INEQ[k] <- index$Gini
     }

dts <- data.frame(INEQ, PTS=Tbox$PTS,
                  CONF=Tadd$Conference)
mypal2 <- colorRampPalette(c("blue","red"))
scatterplot(dts, data.var=c("INEQ","PTS"), z.var="CONF",
            labels=Tbox$Team, palette=mypal,
            repel_labels=TRUE)
```

```{r}
PbP <- PbPmanipulation(PbP.BDB)

PbP.WAS <- subset(PbP, team == "WAS")

lineup <- c("Bradley Beal", "John Wall", "Otto Porter Jr.", "Kelly Oubre Jr.",
             "Marcin Gortat")

filt5 <- apply(PbP.WAS[, 4:13], 1,
               function(x) {
               x <- as.character(x)
               sum(x %in% lineup)==5
               })

subPbP.WAS <- PbP.WAS[filt5, ]

PTS5 <- sapply(lineup,
               function(x) {
               filt <- subPbP.WAS$player==x
               sum(subPbP.WAS$points[filt], na.rm=T)
               })

inequality(PTS5,nplayer=5)
```

```{r}
PbP.WAS.TOR <- subset(PbP, team=="WAS" & oppTeam=="TOR")

filt5.2 <- apply(PbP.WAS.TOR[, 4:13], 1,
               function(x) {
               x <- as.character(x)
               sum(x %in% lineup)==5
               })

subPbP.WAS.TOR <- PbP.WAS.TOR[filt5.2, ]

PTS5 <- sapply(lineup,
               function(x) {
               filt <- subPbP.WAS.TOR$player==x
               sum(subPbP.WAS.TOR$points[filt], na.rm=T)
               })

inequality(PTS5,nplayer=5)
```

## Shot charts
```{r}
#unique(PbP$player)
subdata <- subset(PbP, player=="Stephen Curry")
subdata$xx <- subdata$original_x/10
subdata$yy <- subdata$original_y/10-41.75
```

```{r}
shotchart(data=subdata, x="xx", y="yy", type=NULL,
          scatter=TRUE)
```

```{r}
shotchart(data=subdata, x="xx", y="yy", z="result", type=NULL,
          scatter=TRUE)
```

```{r}
shotchart(data=subdata, x="xx", y="yy", z="playlength", 
          num.sect=5, type="sectors", scatter = TRUE)
```

```{r}
shotchart(data=subdata, x="xx", y="yy", z="playlength", 
          num.sect=5, type="sectors", scatter=FALSE, result="result")
```



# PATTERN IN DATA
## Statistical dependence
```{r}
PbP <- PbPmanipulation(PbP.BDB)

PbP.GSW <- subset(PbP, team=="GSW")

ev <- c("ejection","end of period","jump ball",
        "start of period","unknown","violation",
        "timeout","sub","foul","turnover")

event.unsel <- which(PbP.GSW$event_type %in% ev)

PbP.GSW.ev <- PbP.GSW[-event.unsel,]

attach(PbP.GSW.ev)
T <- table(oppTeam, event_type, exclude=ev)
detach(PbP.GSW.ev)

assocstats(T)
```

## Mean Dependence
```{r}
FF <- fourfactors(Tbox, Obox)
attach(Tbox)
attach(FF)
```

```{r}
df_stats <- data.frame(PTS, P2M, P3M, FTM, REB=OREB+DREB, AST, STL, BLK, ORtg, DRtg)
detach(Tbox)
detach(FF)
```

```{r}
Playoff <- Tadd$Playoff

eta <- sapply(df_stats, function(Y){
  cm <- round(tapply(Y, Playoff, mean), 1)
  eta2 <- etaSquared(aov(Y~Playoff))[1]*100
  c(cm, round(eta2, 2))
}) %>%
  t() %>%
  as.data.frame() %>%
  rename(No=N, Yes=Y, eta2=V3) %>%
  rownames_to_column('rownm') %>%
  arrange(-eta2) %>%
  column_to_rownames('rownm')

eta
```

## Correlation
```{r}
df_ast_tov <- subset(Pbox, MIN>=500)
attach(df_ast_tov)
ast_tov <- data.frame(AST, TOV)/MIN
detach(df_ast_tov)

cor(ast_tov$AST, ast_tov$TOV)
```
ˆ
```{r}
cor(rank(ast_tov$AST), rank(ast_tov$TOV))
```

```{r}
cor(ast_tov$AST, ast_tov$TOV, method="spearman")
```

```{r}
cor(ast_tov)
```

## Linear correlation among variables
```{r}
data <- merge(Pbox, Tadd, by="Team")
data <- subset(data, MIN >= 500)

attach(data)
df_off <- data.frame(
  Player = data$Player,
  PTS = data$PTS / data$MIN,
  P3M = data$P3M / data$MIN,
  P2M = data$P2M / data$MIN,
  REB = (data$OREB + data$DREB) / data$MIN,
  AST = data$AST / data$MIN,
  TOV = data$TOV / data$MIN,
  STL = data$STL / data$MIN,
  BLK = data$BLK / data$MIN,
  Playoff = data$Playoff
)
detach(data)

str(df_off)
```

```{r}
numeric_columns <- sapply(df_off, is.numeric)

corrmatrix <- corranalysis(df_off[,2:9], threshold=0.5)
plot(corrmatrix)
```

```{r}
scatterplot(df_off, data.var=2:9, z.var="Playoff", 
            diag=list(continuous="blankDiag"))
```

```{r}
ggpairs(df_off, columns = 2:9, mapping = aes(color = Playoff)) +
  theme_bw()
```

## Individual cases according to their similarity
```{r}
attach(Pbox)
data2 <- data.frame(PTS, P3M, P2M, REB=OREB+DREB,
                   AST, TOV, STL, BLK)
detach(Pbox)
data2 <- subset(data2, Pbox$MIN>=1500)
id <- Pbox$Player[Pbox$MIN>=1500]
```

```{r}
# Multidimensional scaling
mds <- MDSmap(data2)
```
```{r}
plot(mds, labels=id)
```

```{r}
selp <- which(id=="LeBron James" | id=="Kevin Durant" |
              id=="Jayson Tatum" | id=="Nikola Jokic" |
              id=="Andrew Wiggins")

plot(mds, labels=id, subset=selp, col.subset="red")
```

## Network Relationship
# Cleveland
```{r}
PbP <- PbPmanipulation(PbP.BDB)

PbP.CLE <- subset(PbP, team=="CLE")

netdata <- assistnet(PbP.CLE)
```

```{r}
set.seed(7)
plot(netdata)
```

```{r}
plot(netdata, layout="circle", edge.thr=1)
```

#Golden State Warriors
```{r}
PbP.GSW <- subset(PbP, team=="GSW")

netdataGSW <- assistnet(PbP.GSW)

cols <- paste0(c("a","h"), rep(1:5,each=2))

PbP.GSW.DG0 <- PbP.GSW[!apply(PbP.GSW[,cols], 1, "%in%",
                       x="Draymond Green"),]
netdata.DG0 <- assistnet(PbP.GSW.DG0)

set.seed(1)
plot(netdata.DG0)
```

```{r}
PbP.GSW.DG0 <- subset(PbP.GSW.DG0,
                      ShotType=="2P" | ShotType=="3P")

p0 <- mean(PbP.GSW.DG0$points)
pl0 <- mean(PbP.GSW.DG0$playlength)

PbP.GSW.DG1 <- PbP.GSW[apply(PbP.GSW[,cols], 1, "%in%",
                       x="Draymond Green"),]
PbP.GSW.DG1 <- subset(PbP.GSW.DG1,
                      ShotType=="2P" | ShotType=="3P")

p1 <- mean(PbP.GSW.DG1$points)
pl1 <- mean(PbP.GSW.DG1$playlength)

plot(netdataGSW, layout="circle", edge.thr=20,
     node.col="FGPTS_AST", node.size="ASTPTS")
```

```{r}
plot(netdataGSW, layout="circle", edge.thr=20,
     node.col="FGPTS", node.size="FGPTS_ASTp")
```

```{r}
TAB <- netdataGSW$assistTable
N <- netdataGSW$nodeStats

names(N)[1] <- "Player"

data3 <- merge(N, Pbox, by="Player")

mypal <- colorRampPalette(c("blue","yellow","red"))

scatterplot(data3, data.var=c("FGM","FGM_ASTp"),
            z.var="MIN", labels=data3$Player,
            palette=mypal, repel_labels=TRUE)
```

```{r}
sel <- which(data3$MIN > 984)
print(sel)
dim(TAB)
tab <- TAB[sel, sel]

no.pl <- nrow(tab)

pR <- pM <- vector(no.pl, mode="list")

GiniM <- array(NA, no.pl)
GiniR <- array(NA, no.pl)

for (pl in 1:no.pl) {
     ineqplM <- inequality(tab[pl,], npl=no.pl)
     GiniM[pl] <- ineqplM$Gini
     ineqplR <- inequality(tab[,pl], npl=no.pl)
     GiniR[pl] <- ineqplR$Gini
     title <- rownames(tab)[pl]
     pM[[pl]] <- plot(ineqplM, title=title)
     pR[[pl]] <- plot(ineqplR, title=title)
}
```

```{r}
grid.arrange(grobs=pM, nrow=2)
```

```{r}
grid.arrange(grobs=pR, nrow=2)
```

```{r}
assocstats(tab)
```

```{r}
XX <- data.frame(N[sel,], GiniM, GiniR)

labs <- c("Gini Index for assists made",
          "Gini Index for assists received",
          "Assists received", "Assists made")

bubbleplot(XX, id="Player", x="GiniM", y="GiniR",
           col="FGM_AST", size="AST",
           labels=labs, text.size=4)
```

```{r}
net1 <- as_tbl_graph(netdata$assistNet)
class(net1) <- "igraph"
centr_degree(net1)
```

```{r}
alpha_centrality(net1)
```

```{r}
closeness(net1, mode="all")
```

```{r}
betweenness(net1)
```

## Estimate density events
### Density with respect to a concurrent variable
```{r}
PbP <- PbPmanipulation(PbP.BDB)
unique(PbP$team)

data.team.UTA <- subset(PbP, team=="UTA" & result!="")
data.opp <- subset(PbP, team!="UTA" & result!="")

densityplot(data=data.team.UTA, shot.type="2P",
            var="periodTime", best.scorer=TRUE)
```

```{r}
densityplot(data=data.team.UTA, shot.type="2P",
            var="totalTime", best.scorer=TRUE)
```

```{r}
densityplot(data=data.team.UTA, shot.type="2P",
            var="playlength", best.scorer=TRUE)
```

```{r}
densityplot(data=data.team.UTA, shot.type="2P",
            var="shot_distance", best.scorer=TRUE)
```

```{r}
densityplot(data=data.opp, shot.type="2P",
            var="periodTime", best.scorer=TRUE)
```

```{r}
densityplot(data=data.opp, shot.type="2P",
            var="totalTime",best.scorer=TRUE)
```

```{r}
densityplot(data=data.opp, shot.type="2P",
            var="playlength", best.scorer=TRUE)
```

```{r}
densityplot(data=data.opp, shot.type="2P",
            var="shot_distance", best.scorer=TRUE)
```

```{r}
LJ <- subset(PbP, player=="LeBron James" & result!="")
AD <- subset(PbP, player=="Anthony Davis" & result!="")
JH <- subset(PbP, player=="James Harden" & result!="")
```

```{r}
# James Harden
densityplot(data=JH, shot.type="field",
            var="playlength")

densityplot(data=JH, shot.type="field",
            var="shot_distance")
```

```{r}
# Anthony Davis
densityplot(data=AD, shot.type="field",
            var="playlength")

densityplot(data=AD, shot.type="field",
            var="shot_distance")
```

```{r}
# LeBron James
densityplot(data=LJ, shot.type="field",
            var="playlength")

densityplot(data=LJ, shot.type="field",
            var="shot_distance")
```

## Density in Space
```{r}
PbP <- PbPmanipulation(PbP.BDB)

PbP$xx <- PbP$original_x/10
PbP$yy <- PbP$original_y/10 - 41.75

LJ <- subset(PbP, player=="LeBron James")

shotchart(data=LJ, x="xx", y="yy",
          type="density-polygons")
```

```{r}
shotchart(data=LJ, x="xx", y="yy", type="density-raster")
```

```{r}
shotchart(data=LJ, x="xx", y="yy", type="density-hexbin")
```

```{r}
shotchart(data=LJ, x="xx", y="yy",
          type="density-polygons", scatter=TRUE)
```

```{r}
shotchart(data=LJ, x="xx", y="yy", type="density-raster",
          scatter=TRUE, pt.col="tomato", pt.alpha=0.1)
```

```{r}
shotchart(data=LJ, x="xx", y="yy", type="density-hexbin",
          nbins=50, palette="bwr")
```

### Joint density of two variables
```{r}
data.density <- subset(Pbox, MIN>=500)

attach(data.density)
X.density <- data.frame(PTS, P3M, P2M, REB=OREB+DREB, AST)/MIN
detach(data.density)

scatterplot(X.density, data.var=1:5,
            lower=list(continuous="density"),
            diag=list(continuous="densityDiag"))
```

# FINDING GROUP IN DATA
## Cluster Analysis
```{r}
FF <- fourfactors(Tbox,Obox)

OD.Rtg <- FF$ORtg/FF$DRtg

F1.r <- FF$F1.Off/FF$F1.Def

F2.r <- FF$F2.Def/FF$F2.Off

F3.Off <- FF$F3.Off

F3.Def <- FF$F3.Def

P3M <- Tbox$P3M

STL.r <- Tbox$STL/Obox$STL

data <- data.frame(OD.Rtg, F1.r, F2.r, F3.Off, F3.Def,
                   P3M, STL.r)
```

```{r}
set.seed(123)

kclu1 <- kclustering(data)

plot(kclu1)
```

```{r}
set.seed(123)

kclu2 <- kclustering(data, labels=Tbox$Team, k=5)

plot(kclu2)
```

```{r}
kclu2.PO <- table(kclu2$Subjects$Cluster, Tadd$Playoff)

kclu2.W <- tapply(Tbox$W, kclu2$Subjects$Cluster, mean)

Xbar <- data.frame(cluster=c(1:5), N=kclu2.PO[,1],
                   Y=kclu2.PO[,2], W=kclu2.W)

barline(data=Xbar, id="cluster", bars=c("N","Y"),
        labels.bars=c("Playoff: NO","Playoff: YES"),
        line="W", label.line="average wins",
        decreasing=FALSE)

ggplot(data = Xbar, aes(x = cluster)) +
  geom_bar(aes(y = N, fill = "Playoff: NO"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = Y, fill = "Playoff: YES"), stat = "identity", position = "dodge") +
  geom_line(aes(y = W * 10, group = 1, color = "average wins"), size = 1) +
  geom_text(aes(y = N + 10, label = N), vjust = -0.2, size = 3, position = position_dodge(width = 0.9)) +
  geom_text(aes(y = Y + 10, label = Y), vjust = -0.5, size = 6, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Playoff: NO" = "red", "Playoff: YES" = "green")) +
  scale_color_manual(values = c("average wins" = "blue")) +
  labs(
    x = "Cluster",
    y = "Count",
    fill = "",
    color = "",
    title = "Barline Chart with Labels",
    subtitle = "Comparison of No, Yes, and Average Wins by Cluster"
  ) +
  theme_minimal()
# The number with a bigger size are for "Yes' and smaller size are for "No"
```

```{r}
cluster <- as.factor(kclu2$Subjects$Cluster)

Xbubble <- data.frame(Team=Tbox$Team, PTS=Tbox$PTS,
                      PTS.Opp=Obox$PTS, cluster,
                      W=Tbox$W)

labs <- c("PTS", "PTS.Opp", "cluster", "Wins")

bubbleplot(Xbubble, id="Team", x="PTS", y="PTS.Opp",
           col="cluster", size="W", labels=labs)
```

## K-means clustering of Toronto Raptors shots
```{r}
PbP <- PbPmanipulation(PbP.BDB)

shots.TOR <- subset(PbP,
                !is.na(PbP$shot_distance) &
                PbP$team=="TOR")
shots.TOR <- mutate_if(shots.TOR, is.factor, droplevels)

attach(shots.TOR)
data.kclu <- data.frame(PTS=points, DIST=shot_distance,
                   TIMEQ=periodTime, PL=playlength)
detach(shots.TOR)

set.seed(1)
kclu.TOR <- kclustering(data.kclu, algorithm="MacQueen",
                     nclumax=15, iter.max=500)
plot(kclu.TOR)
```

```{r}
set.seed(246)

kclu.TOR2 <- kclustering(data.kclu, algorithm="MacQueen",
                     iter.max=500, k=6)

plot(kclu.TOR2)
```

```{r}
cluster <- as.factor(kclu.TOR2$Subjects$Cluster)
shots.TOR <- data.frame(shots.TOR, cluster)
shots.TOR$xx <- shots.TOR$original_x/10
shots.TOR$yy <- shots.TOR$original_y/10 - 41.75

no.clu.TOR <- 6

p1 <- p2 <- vector(no.clu.TOR, mode="list")

for (k in 1:no.clu.TOR) {
     shots.k <- subset(shots.TOR,cluster==k)
     p1[[k]] <- shotchart(data=shots.k, x="xx", y="yy",
                          z="result", type=NULL,
                          scatter = TRUE,
                          drop.levels=FALSE)
p2[[k]] <- shotchart(data=shots.k, x="xx", y="yy",
                     z="periodTime",
                     col.limits=c(0,720),
                     result="result", num.sect=5,
                     type="sectors", scatter=FALSE)
     }
```

```{r}
grid.arrange(grobs=p1, nrow=3)
```

```{r}
grid.arrange(grobs=p2, nrow=3)
```

```{r}
shots.pl <- table(shots.TOR$player, shots.TOR$cluster)

Xineq <- as.data.frame.matrix(shots.pl)

no.clu.TOR <- 6

p <- vector(no.clu.TOR, mode="list")

for (k in 1:no.clu.TOR) {
     ineqC <- inequality(Xineq[,k], npl=nrow(Xineq))
     title <- paste("Cluster", k)
     p[[k]] <- plot(ineqC, title=title)
}

grid.arrange(grobs=p, nrow=3)
```

```{r}
shots.perc <- shots.pl/rowSums(shots.pl)

Xbar.TOR <- data.frame(player=rownames(shots.pl),
                   rbind(shots.perc),
                   FGA=rowSums(shots.pl))

labclusters <- c("Cluster 1","Cluster 2","Cluster 3",
                 "Cluster 4","Cluster 5","Cluster 6")

barline(data=Xbar.TOR, id="player", line="FGA",
        bars=c("X1","X2","X3","X4","X5","X6"),
        order.by="FGA", label.line="Field goals attempted",
        labels.bars=labclusters)

ggplot(Xbar.TOR, aes(x=player, y=FGA)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=X1), vjust=-0.3, size=3.5) +
  theme_minimal()

ggplot(Xbar.TOR, aes(x=player, y=X1)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=X1), vjust=-0.3, size=3.5) +
  theme_minimal()

ggplot(Xbar.TOR, aes(x=player, y=X2)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=X1), vjust=-0.3, size=3.5) +
  theme_minimal()

ggplot(Xbar.TOR, aes(x=player, y=X3)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=X1), vjust=-0.3, size=3.5) +
  theme_minimal()

ggplot(Xbar.TOR, aes(x=player, y=X4)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=X1), vjust=-0.3, size=3.5) +
  theme_minimal()

ggplot(Xbar.TOR, aes(x=player, y=X5)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=X1), vjust=-0.3, size=3.5) +
  theme_minimal()

ggplot(Xbar.TOR, aes(x=player, y=X6)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=X1), vjust=-0.3, size=3.5) +
  theme_minimal()
```

## Hierarchical clustering of NBA players
```{r}
rm(list=ls())

attach(Pbox)
df <- data.frame(PTS, P3M, REB=OREB+DREB,
                   AST, TOV, STL, BLK, PF)
detach(Pbox)

df <- subset(df, Pbox$MIN>=1500)

ID <- Pbox$Player[Pbox$MIN>=1500]

hclu <- hclustering(df)
plot(hclu)
```

```{r}
hclu2 <- hclustering(df, labels=ID, k=9)

plot(hclu2, profiles=TRUE)
```

```{r}
plot(hclu2, rect=TRUE, colored.branches=TRUE,
     cex.labels=0.5)
```

```{r}
Pbox.subset <- subset(Pbox, MIN>=1500)

MIN <- Pbox.subset$MIN

X <- data.frame(hclu2$Subjects, scale(df), MIN)

dvar <- c("PTS","P3M","REB","AST",
          "TOV","STL","BLK","PF")
svar <- "MIN"

yRange <- range(X[,dvar])

sizeRange <- c(1500, 3300)

no.clu <- 9

p <- vector(no.clu, mode="list")

for (k in 1:no.clu) {
     XC <- subset(X, Cluster==k)
     vrb <- variability(XC[,3:11], data.var=dvar,
                        size.var=svar, weight=FALSE,
                        VC=FALSE)
     title <- paste("Cluster", k)
     p[[k]] <- plot(vrb, size.lim=sizeRange, ylim=yRange,
               title=title, leg.pos=c(0,1),
               leg.just=c(-0.5,0),
               leg.box="vertical",
               leg.brk=seq(1500,3000,500),
               leg.title.pos="left", leg.nrow=1,
               max.circle=7)
     }

grid.arrange(grobs=p, ncol=3)
```

# LINEAR MODELS
## Simple linear regression
```{r}
rm(list=ls())

Pbox.sel <- subset(Pbox, MIN>=500)

attach(Pbox.sel)
x <- AST/MIN
y <- TOV/MIN
pl <- Player
detach(Pbox.sel)

out <- simplereg(x, y, type="lin")
xtitle <- "AST per minute"
ytitle <- "TOV per minute"
plot(out, xtitle=xtitle, ytitle=ytitle)
```

```{r}
selp <- which(pl=="Kyrie Irving")
plot(out, labels=pl, subset=selp, xtitle=xtitle,
     ytitle=ytitle)
```

```{r}
plot(out, labels=pl, subset="quant",
     Lx=0, Ux=0.97, Ly=0, Uy=0.97,
     xtitle=xtitle, ytitle=ytitle)
```

## Non Parametric Regression
### Polynomial regression
```{r}
rm(list=ls())

Pbox.sel <- subset(Pbox, MIN>=500)

attach(Pbox.sel)
x <- (DREB+OREB)/MIN
y <- P3M/MIN
pl <- Player
detach(Pbox.sel)

out <- simplereg(x, y, type="lin")
xtitle <- "REB per minute"
ytitle <- "P3M per minute"
plot(out, xtitle=xtitle, ytitle=ytitle)
```

```{r}
out <- simplereg(x, y, type="pol")
plot(out, labels=pl, subset="quant",
     Lx=0, Ux=0.90, Ly=0, Uy=0.95,
     xtitle=xtitle, ytitle=ytitle)
```

### Gaussian kernel smoothing
```{r}
data <- subset(Pbox, MIN>=500)

attach(data)
df <- data.frame(PTS, P3M, P2M, REB=OREB+DREB, AST)/MIN
detach(data)

scatterplot(df, data.var=1:5,
lower=list(continuous="smooth_loess"),
diag=list(continuous="barDiag"))
```

###  Estimation of scoring probability
```{r}
PbP <- PbPmanipulation(PbP.BDB)

PbP.GSW <- subset(PbP, team=="GSW" & result!="")

p1 <- scoringprob(data=PbP.GSW, shot.type="3P",
                  var="playlength")

p2 <- scoringprob(data=PbP.GSW, shot.type="3P",
                  var="periodTime", bw=300)

grid.arrange(p1, p2, ncol=2)
```

```{r}
pl1 <- c("Kevin Durant","Stephen Curry","Klay Thompson")

p1 <- scoringprob(data=PbP.GSW, shot.type="2P",
players=pl1, var="shot_distance",
col.team="gray")

pl2 <- c("Kevin Durant","Stephen Curry")

p2 <- scoringprob(data=PbP.GSW, shot.type="2P",
players=pl2, var="totalTime", bw=1500,
col.team="gray")

grid.arrange(p1, p2, ncol=2)
```

### Estimation of expected points
```{r}
mypal <- colorRampPalette(c("red","green"))

expectedpts(data=PbP.GSW, players=pl1,
            col.team="gray", palette=mypal,
            col.hline="gray")
```

```{r}
Pbox.GSW <- subset(Pbox, PTS>=500 &
                   Team=="Golden State Warriors")

pl.gsw <- Pbox.GSW$Player

mypal <- colorRampPalette(c("red","green"))

expectedpts(data=PbP.GSW, players=pl.gsw,
            col.team="gray", palette=mypal,
            col.hline="gray")
```

```{r}
expectedpts(data=PbP.GSW, bw=300, players=pl.gsw,
            col.team="gray", palette=mypal,
            col.hline="gray", var="periodTime",
            xlab="Period time")
```

```{r}
top <- subset(Tadd, Playoff=="Y" & team!="GSW")$team
bot <- subset(Tadd, Playoff=="N")$team 
 
bot_top <- function(X, k) {
           dts <- subset(subset(X, oppTeam %in% get(k)),
                         team=="GSW")
           dts$player <- paste(dts$player, k)
           return(dts)
           }

PbP.GSW2 <- rbind(bot_top(PbP, "top"), bot_top(PbP, "bot"))

pl.gsw2 <- c("Stephen Curry top","Stephen Curry bot",
        "Kevin Durant top", "Kevin Durant bot")

mypal <- colorRampPalette(c("red","green"))

expectedpts(data=PbP.GSW2, bw=1200, players=pl.gsw2,
            col.team="gray", palette=mypal,
            col.hline="gray", var="totalTime",
            xlab="Total time", x.range=NULL)
```

