---
title: "Fish Market"
author: "Karl Mbouombouo"
date: "2023-08-21"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
library(readr)
library("ggplot2")
library("dplyr")
library("broom")
library("ggpubr")
library("psych")
library('ggiraphExtra')
```

## Data

```{r fish}
fish.data <- read_csv("~/Desktop/R-Projects/Multiple Linear Regression/Fish Market/Fish.csv")

which(is.na(fish.data))

summary(fish.data)

dim(fish.data)
```

## Correlation and Plots

```{r pressure, echo=FALSE}
corPlot(fish.data[3:7], cex = 0.8)
```
An highly correlation exist between Length1, Length2, and Length3, so we cannot 
include those parameters together in the model.

```{r}
# Relationship between Weight and Length1
ggplot(fish.data, aes(x = Length1, y = Weight)) +
  geom_point()

# Relationship between Weight and Height
ggplot(fish.data, aes(x = Height, y = Weight)) +
  geom_point()

# Relationship between Weight and Width
ggplot(fish.data, aes(x = Width, y = Weight)) +
  geom_point()
```

```{r}
hist(fish.data$Weight)
```
The distribution of observations is roughly right-skewed, the mean is greater 
than the median.

## Modeling
```{r}
fit <- lm(Weight ~ Length1 + Height + Width, data = fish.data)
summary(fit)
```
The estimated effect of the length1 on weight is 22.683, while the estimated 
effect of height is 13.670, and width is 44.069.
The p-value here it's 2.2e-16 (or almost zero), which will indicate whether the 
model fits the data well.
We can say that there is a significant positive relationship between Weight and 
Lenght1, Height and Width, with respectively 22.683 unit, 13.670 unit and 44.069
increase in fish's weight for every unit increase in Lenght1, Height and Width.

## Result visualization
```{r}
par(mfrow=c(2,2))
plot(fit)
par(mfrow=c(1,1))
```
The residuals are all basically horizontal and centered around zero. This means
there are no outliers or biases in the data that would make a linear regression 
invalid.
From the Normal Q-Q plot in the top right, we can see that the real residuals 
from our model form an almost perfectly one-to-one line with the theoretical 
residuals from a perfect model.
Based on these residuals, we can say that our model meets the assumption of 
homoscedasticity.

## Conclusion
In our fish surveys, we found significant relationships between the fish weight 
and the length, height and width (p < 0, p < 0.001, and p < 0.005 respectively).
Specifically we found that 44.069 increase (± 15.347) in weight (gramme) for 
every 1cm increase in width (centimeter), 22.683 increase (± 2.021) in weight
for every 1cm increase in length, and 13.670 increase (± 3.854) in weight
for every 1cm increase in width.
