---
title: "Denver Broncos - Assessment"
author: "Karl Mbouombouo"
date: "2023-07-28"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
library(dlookr)
```

## Explore & Clean Dataset

```{r, results='hide'}
procedures <- read_csv("procedures.csv")
procedure_details <- read_csv("procedure_details.csv")
pets <- read_csv("pets.csv")
owners <- read_csv("owners.csv")

#Missing Values (NA)
which(is.na(procedures))
which(is.na(procedure_details))
which(is.na(pets))
which(is.na(owners))
```

# What is the name of the oldest dog in Southfield?
```{r, warning=FALSE}
southfield_pets <- inner_join(owners, pets, by = "OwnerID")

# Filter the data to include only pets in Southfield
southfield_pets <- southfield_pets %>%
  filter(City == "Southfield" & Kind == "Dog")

# Find the oldest dog in Southfield
oldest_dog <- southfield_pets %>%
  filter(Age == max(Age))

# Extract the name of the oldest dog
oldest_dog_name <- oldest_dog$Name.y

# Print the name of the oldest dog
print(oldest_dog_name)
```
The name of the oldest dog is Crockett

# What is the average (mean) number of pets per city?
```{r, warning=FALSE}
owners_pets <- merge(owners, pets, by = "OwnerID", all.x=TRUE, all.y=FALSE)

# Count the number of pets in each city
pets_per_city <- owners_pets %>%
  group_by(City) %>%
  summarise(total_pets = n())

# Calculate the average number of pets per city
average_pets_per_city <- mean(pets_per_city$total_pets)

# Print the result
print(average_pets_per_city)
```
2.5

# Which owner spent the most on procedures for their pet(s)?
```{r, warning=FALSE}
# Combine the data
df_2 <- inner_join(owners, pets, by = "OwnerID") %>%
  inner_join(procedures, by = "PetID") %>%
  inner_join(procedure_details, by = "ProcedureSubCode")

# Calculate the total amount spent on procedures for each owner
total_spent_per_owner <- df_2 %>%
  group_by(OwnerID, Name.x, Surname) %>%
  summarise(total_spent = sum(Price))
            
# Find the owner who spent the most on procedures
owner_with_max_spending <- total_spent_per_owner %>%
  filter(total_spent == max(total_spent))

# Print the result
print(owner_with_max_spending)
```
Lee Mckenzie

# How many owners spent 20 dollars or more on procedures for their pets?
```{r, warning=FALSE}
# Calculate the total cost of procedures for each pet based on 'ProcedureType' and 'ProcedureSubCode'
total_cost_per_pet <- procedure_details %>%
  left_join(procedures, by = c("ProcedureType", "ProcedureSubCode")) %>%
  group_by(PetID) %>%
  summarize(TotalCost = sum(Price, na.rm = TRUE))

# Merge the 'pets' and 'total_cost_per_pet' data on PetID
merge_1 <- merge(pets, total_cost_per_pet, by = "PetID")

# Calculate the total procedure cost per owner
total_cost_per_owner <- merge_1 %>%
  group_by(OwnerID) %>%
  summarize(TotalCost = sum(TotalCost))

# Filter and count the number of owners who spent 20 dollars or more on procedures
spent_20_or_more <- total_cost_per_owner %>%
  filter(TotalCost >= 20) %>%
  nrow

print(spent_20_or_more)
```
12

# How many owners have at least two different kinds of pets?
```{r, warning=FALSE}
# Group by 'OwnerID' and count the distinct number of 'Kind' values for each owner
multiple_kinds <- pets %>%
  group_by(OwnerID) %>%
  summarize(NumDifferentKinds = n_distinct(Kind))

# Filter the data to include only the owners with at least two different kinds of pets
owners_with_multiple_kinds <- multiple_kinds %>%
  filter(NumDifferentKinds >= 2) %>%
  nrow()

# Print the result
print(owners_with_multiple_kinds)
```
4

# How many owners have pets where the first letter of their name (OwnerName) matches the first letter of their pet's name (PetName)? 
```{r, warning=FALSE}
df_3 <- inner_join(owners, pets, by = "OwnerID")

# Count the number of owners where the first letter of their name matches the first letter of their pet's name
matching_first_letter <- df_3 %>%
  filter(substr(Name.x, 1, 1) == substr(Name.y, 1, 1)) %>%
  summarise(num_owners = n_distinct(OwnerID))

# Print the result
print(matching_first_letter$num_owners)
```
8

# What percentage of cities have more male pets than female pets
```{r, warning=FALSE}
df_4 <- inner_join(pets, owners, by = "OwnerID")

# Calculate the total number of male pets and female pets in each city
pet_count_by_city <- df_4 %>%
  group_by(City) %>%
  summarise(total_male_pets = sum(Gender == "male"),
            total_female_pets = sum(Gender == "female"))

# Count the number of cities where the total number of male pets is greater than the total number of female pets
cities_with_more_male_pets <- pet_count_by_city %>%
  filter(total_male_pets > total_female_pets) %>%
  nrow()

# Calculate the percentage of cities with more male pets than female pets
total_cities <- nrow(pet_count_by_city)
percentage_more_male_pets <- (cities_with_more_male_pets / total_cities) * 100

# Print the result
print(percentage_more_male_pets)
```
57.5

# What percentage of pets received a vaccination?
```{r, warning=FALSE}
df_5 <- inner_join(pets, procedures, by = "PetID")

# Count the number of pets that received a vaccination
pets_with_vaccination <- df_5 %>%
  filter(ProcedureType == "VACCINATIONS") %>%
  nrow()

# Calculate the percentage of pets that received a vaccination
total_pets <- nrow(pets)
percentage_vaccinated <- (pets_with_vaccination / total_pets) * 100

# Print the result
print(percentage_vaccinated)
```
30% (Approximation)


# Which city's pet sample is made up of exactly 70% dogs? The answer is case sensitive, so please match the value for City exactly.
```{r, warning=FALSE}
# Combine data from 'table 3' and 'table 4' based on 'OwnerID'
df_6 <- pets %>%
  left_join(owners, by = "OwnerID")

# Filter rows where 'Kind' is 'Dog', group by 'City', and count the number of dogs in each city
# Group by 'City' and calculate the total number of pets in each city
result <- df_6 %>%
  group_by(City) %>%
  summarize(Num_Dogs = sum(Kind == "Dog"),
            Total_Pets = n())

# Calculate the ratio of dogs to total pets
result$Dog_Ratio <- (result$Num_Dogs / result$Total_Pets) * 100

# Display the table
filtered_70<- filter(result, Dog_Ratio == 70)
print(filtered_70)
```
Grand Rapids
